Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-15 paper.tex
word change
final changes

\documentclass{IEEEtran}

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    Big Data processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype, as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality, with a number of concurrently processed subject in the  280 ballpark. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines reduce data-related overheads through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, executes tasks in the nodes holding the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, enables further Big Data performance 
optimizations, by evaluating results only when necessary. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

@@

Our study focuses on performance. We intentionally do not 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture 
and representation, portability or reproducibility which are otherwise 
critical concerns, addressed for instance in~\cite{samba}. Besides, our 
study of performance focuses on the impact of data writes and 
transfers. It purposely leaves out task scheduling to computing 
resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion, however.

@@

Apache Spark is a popular Scala-based processing framework for Big Data, with
APIs in Java, Python and R. Its 
generalized nature allows Spark to not only be applied to batch workflows,
but also SQL queries, iterative machine learning applications, 
data streaming and graph processing. Spark's
main features include data-locality, in-memory processing and lazy evaluation,
which it achieves through its principle abstraction, the Resilient Distributed 
Dataset (RDD).

@@

All actions and wide dependencies require a shuffle -- Spark's most costly
operation. Every shuffle begins with each map task saving its data to local
files for fault tolerance. The shuffle operation then 
redistributes the data across the partitions as requested. A shuffle marks a 
stage boundary in Spark. In other words, an reduce operation  will not begin 
until all 
dependent map tasks have completed.

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

Data locality in Spark is achieved through the scheduling of tasks to partitions 
which have the
data loaded in memory. If the data is instead stored on HDFS, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.
%improve paragraph

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
clusters and is prepackaged with Hadoop installations. In contrast to YARN, 
Mesos was designed to be used in multitenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster.

@@

% i use 'popular' for both... widely-used (for Spark)?
Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as the SunGrid Engine (SGE), Portable Batch System (PBS), Slurm and HTCondor\todo{cite}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype also includes many built-in 
interfaces to commonly used neuroimaging tools to be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modification to achieve this.

Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
the job will execute on a single input. However, the MapNode, 
a child variant of the Node, can execute on multiple inputs. All tasks in
Nipype execute in their own uniquely named subdirectory which facilitates provenance 
tracking of inputs and outputs, and also enables checkpointing of the workflow.
In the case of Node failure or application modification, only the nodes who have
been modified (verified by hash) are re-executed.

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared 
local filesystem, but it may come at the expense of fault-tolerance, as 
a result of being stored in a disk-local file system \todo{I don't get 'as a result of being stored in a disk-local file system'}. Moreover, the 
user will need to ensure that the files are appropriately directed to 
the nodes that require them as there is no guarantee of data locality 
in Nipype.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose content reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. In case engine memory is used, the engine must cleanup 
 unused data with some strategy to avoid crashes. In case an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

@@

Storing data on local disks obviously enables data locality as tasks 
executed on the node where their input files reside do not require to 
transfer input or output data. However, in absence of a more specific 
file system such as HDFS to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
 A 698TB Lustre server was accessible on each node through a 100Gb/s 
 Infiniband network. The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

As it is possible that both small and large data may need to be processed using
the same analysis pipeline, we 
wanted to determine the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to 
subject 1 of OpenNeuro's 
\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
dataset version 6}. In order to be able to split it into 125 chunks, it 
was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

@@

In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would be read from / written to  Lustre. 
We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality \todo{You could refer to the script doing that in the paper repo.}. That is, the chunk were split into partitions, and for each partition of 
chunks, we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

For our first incrementation experiment, we investigated the effects of 
Big Data strategies on varying total data size. To do this, we 
increased the number of incrementation iterations from 1, 10 and 100 
times. The total data size would then increase from 75GB, at 1 
iteration, to 7,500GB, at 100 iterations. The total number of chunks 
was 125. Chunks all ran concurrently ($\Gamma$=125) and they were 
equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
node ($\gamma$=9). Task duration was fixed at 3.44 seconds.

In the second experiment, we evaluated the effects of Big Data 
strategies on varying task duration. If the page cache has sufficient 
time to flush, it should be expected that using in-memory computing 
should have no increased speed-up over disk. We varied the task 
duration between 2.4 seconds and 320 seconds so that the D/C falls into 
different regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}. The number of chunks was maintained 
at 125, leading to $\Gamma$=125 and $\gamma$=9. The number of iterations was
fixed to 10.

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, greater chunk size signifies 
a decrease in parallelization, however, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. Unfortunately, 30 chunks could 
no be processed using in-memory Spark as its individual chunk size of 
3GB exceeded the 2GB partition size limit imposed by Spark \todo{You should move this comment to the results or the reader will wonder
what happened to the yellow bar when looking at figure 1.c.}. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed at 125. Similarly to the previous experiment, the 
total sequential compute time was fixed (10 iterations, 1.76 seconds 
per task), however, due to varying size in total data processed (D), 
the D/C ratio varied. Once again, we ensured that the D/C ratio fell in 
multiple different regions of the inequality. The D/C ratio ranged from 
349.1 MB/s for BigBrain and 174.6MB/s for half Bigbrain, to 0.06MB/s for the MRI 
image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
disk.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the page cache flushed. \todo{Update the graph to the same y scale and 'broken'
bars, and check that this impression still holds...}
This is explained by the fact that the application data was not large enough to saturate the page  
cache fully (transient phase). As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

%sp_mem 1it = 34s, 10it = 64 100it = 381
%3.44 * 1 = 3.44, 3.44*10 = 34, 3.44*100 = 344

%sp_tmpfs 1it = 31, 10it = 69,

At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82~x 
slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
total memory. Since there was 187~GB of RAM on each node, only 37.4~GB of 
dirty data could be held in a node's page cache at any given time. The 
slowdown experienced can be explained by the smaller percentage of total data 
residing in the page cache at a given time, compared to 10 iterations. 
Therefore, the effects of Lustre bandwidth are more significant in 
this application. At 100 iterations, the application was writing 500GB 
per node (7,500 GB / 15 nodes) and hence could not run on tmpfs or local disk.

@@

% Comment on the general trend
% note 'on average' = average between spark and nipype. local disk spark is 
% 1.22x slower and local disk nipype is 1.54x slower
% spark lustre 2.66x slower; nipype lustre 2.68x slower

>>>>

\documentclass{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Lazy evaluation strategies
were found to increase the likelihood of cache hits, further improving processing time. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality and processing around 280 subjects concurrently. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines reduce data-related overheads through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, execute tasks within the nodes which store the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, enables further Big Data performance 
optimizations, by evaluating results only when necessary. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

@@

Our study focuses on performance. We intentionally do not 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture 
and representation, portability or reproducibility, which are otherwise 
critical concerns, addressed for instance in~\cite{samba}. Besides, our 
study of performance focuses on the impact of data writes and 
transfers. It purposely leaves out task scheduling to computing 
resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion, however.

@@

Apache Spark is a well-established Scala-based processing framework for Big Data, with
APIs in Java, Python and R. Its 
generalized nature allows Spark to not only be applied to batch workflows,
but also SQL queries, iterative machine learning applications, 
data streaming and graph processing. Spark's
main features include data-locality, in-memory processing and lazy evaluation,
which it achieves through its principle abstraction, the Resilient Distributed 
Dataset (RDD).

@@

All actions and wide dependencies require a shuffle -- Spark's most costly
operation. Every shuffle begins with each map task saving its data to local
files for fault tolerance. The shuffle operation then 
redistributes the data across the partitions as requested. A shuffle marks a 
stage boundary in Spark, where reduce-like operations will not begin 
until all dependent map tasks have completed.

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur within a single application, Spark implements a
Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

Data locality in Spark is achieved through the scheduling of tasks to partitions 
which have the
data loaded in memory. If the data is instead stored on HDFS, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler, whereas YARN is scheduler designed for Hadoop~\cite{white2012hadoop}
clusters and is prepackaged with Hadoop installations. In contrast to Standalone and YARN, 
Mesos was designed to be used in multi-tenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster.

@@

% i use 'popular' for both... widely-used (for Spark)?
Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as the Sun/Oracle Grid Engine (SGE/OGE), \href{http://www.adaptivecomputing.com/products/torque/}{TORQUE},
\href{https://slurm.schedmd.com/}{Slurm} and \href{https://research.cs.wisc.edu/htcondor/}{HTCondor}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype additionally provides many built-in 
interfaces to commonly used neuroimaging tools that can be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modifications to achieve this.

Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
the job will execute on a single input. However, the MapNode, 
a child variant of the Node, can execute on multiple inputs. All tasks in
Nipype execute in their own uniquely named subdirectory which facilitates provenance 
tracking of inputs and outputs, and also enables checkpointing of the workflow.
In the case of Node failure or application modification, only the nodes which have
been modified (verified by hash) or have no successfully completed, are re-executed.

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared 
local filesystem, but it may come at the expense of fault-tolerance, as data 
located on failed nodes will be permanently lost. Moreover, the 
user will need to ensure that the files are appropriately directed to 
the nodes that require them as there is no guarantee of data locality 
in Nipype.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose contents reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. When engine memory is used, the engine must cleanup 
 unused data with some strategy to avoid crashes. When an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

@@

Storing data on local disks obviously enables data locality, as tasks 
executed on the node where their input files reside do not require 
transfer of input or output data. However, in absence of a more specific 
file system such as HDFS to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes.
 Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
 as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
 (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
 Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
 memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
 local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
 accesible on each compute node through a 100~GB/s Intel OmniPath network. All
 the nodes connect to a director switch in a 1:1 non-blocking topology.
 The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

Since both small and large datasets may need to be processed using
the same analysis pipeline, we 
examined the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to 
subject 1 of OpenNeuro's 
\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
dataset version 6}. In order to be able to split it into 125 chunks, it 
was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

@@

In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is not a time consuming process,
we added a sleep delay to the incrementation stage to study the effects 
of tasks duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same file system. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would reside on Lustre. 
We chose to perform our initial input/final output on Lustre as local storage is
typically only accessible to a user for the duration of the execution in HPC
environments.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark\_scripts/run\_benchmarks.py}{run\_benchmarks.py}).
That is, the chunk were split into partitions, and for each partition,
we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

For our first incrementation experiment, we investigated the effects of 
Big Data strategies on varying total data size. To achieve this, we 
increased the number of incrementation iterations from 1, 10 and 100 
times. The total data size would then increase from 75GB, at 1 
iteration, to 7,500GB, at 100 iterations. The total number of chunks 
was 125. Chunks all ran concurrently ($\Gamma$=125) and they were 
equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
node ($\gamma$=9). Task duration was fixed at 3.44 seconds.

In the second experiment, we evaluated the effects of Big Data 
strategies on varying task duration. If the page cache has sufficient 
time to flush, it would be expected that in-memory computing 
and local disk perform equivalently. We varied the task 
duration between 2.4 and 320 seconds such that the D/C falls into 
different regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}. The number of chunks was maintained 
at 125, leading to $\Gamma$=125 and $\gamma$=9. The number of iterations was
fixed to 10.

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
a decrease in parallelization. However, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed to 125. Similarly to the previous experiment, the 
total sequential compute time was fixed (10 iterations, 1.76 seconds 
per task), however, due to varying size in total data processed (D), 
the D/C ratio varied. Once again, we ensured that the D/C ratio fell in 
multiple different regions of the inequality. The D/C ratio ranged from 
349.1 MB/s for BigBrain and 174.6MB/s for half Bigbrain, to 0.06MB/s for the MRI 
image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
disk.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the
disk bandwidth. This is because application data was not saturating the Page Cache
(transient phase). The page cache, on Zenith, occupies 40\% of 
total memory. With 192~GB of RAM on each node, 76.8~GB of 
dirty data could be held in a node's page cache at any given time.
As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82~x 
slower than Spark in-memory. The 
slowdown experienced can be explained by the smaller percentage of total data 
residing in the page cache at a given time, compared to 10 iterations. 
Therefore, the effects of Lustre bandwidth are more significant in 
this application. At 100 iterations, the application was writing 500GB 
per node (7,500 GB / 15 nodes) and hence could not run on tmpfs or local disk.

@@

% Comment on the general trend
